---
title: "R Lab"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
urlcolor: "blue"
params:
  solutions: FALSE
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = params$solutions, 
                      eval = params$solutions, 
                      fig.align  = "center",
                      fig.height = 3)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

# County Demographic Information (CDI)

You can read about the CDI dataset [here](https://dcgerard.github.io/stat_415_615/data.html#County_Demographic_Information), and download it from here: <https://dcgerard.github.io/stat_415_615/data/cdi.csv>

Complete this lab by writing solutions in an R markdown document, knitting to word or PDF.

1. Load the CDI dataset into R, saving the data frame in a variable called `cdi`.

    ```{r}
    library(tidyverse)
    cdi <- read_csv("https://dcgerard.github.io/stat_415_615/data/cdi.csv")
    ```

2. Use a function that provides a rough glimpse at these data.

    ```{r}
    glimpse(cdi)
    ```

3. Calculate the mean and standard deviation of the population of the counties.

    ```{r}
    mean(cdi$pop)
    sd(cdi$pop)
    ```

4. Convert the `pop` variable to be in units of 100,000 individuals (so 1 corresponds to 100,000, 2 corresponds to 200,000, etc).

    ```{r}
    cdi <- mutate(cdi, pop = pop / 100000)
    ```

5. Use an appropriate plot to explore the association between population and number of physicians.

    ```{r}
    ggplot(data = cdi, mapping = aes(x = pop, y = physicians)) +
      geom_point()
    ```

6. Use an appropriate plot to explore the association between region and population.

    ```{r}
    ggplot(data = cdi, mapping = aes(x = region, y = pop)) +
      geom_boxplot()
    ```

7. Use an appropriate plot to explore the association between population and number of physicians in each region.

    ```{r}
    ggplot(data = cdi, mapping = aes(x = pop, y = physicians)) +
      geom_point() +
      facet_wrap( . ~region)
    ```

8. Create four datasets, one for each region.

    ```{r}
    cdi_nc <- filter(cdi, region == "NC")
    cdi_ne <- filter(cdi, region == "NE")
    cdi_s <- filter(cdi, region == "S")
    cdi_w <- filter(cdi, region == "W")
    ```

9. Reproduce the following using markdown:

    1. It was the best of times, 
    2. it was the worst of times, 
    - **it was the age of wisdom**, 
    - *it was the age of foolishness,* 
    - [it was the epoch of belief](https://www.youtube.com/watch?v=dQw4w9WgXcQ) ...

