---
title: "Final Project"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
bibliography: "source.bib"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Important Dates

- 11/02/2022: Journal article selected.
- 11/16/2022: Progress report due.
- 12/14/2022: 
    - Final project presentations. 
    - Written report due.
    - Code due.

# General Description

- For the final project you will work in teams of 2--3 to select a journal article from 2022 and reproduce their results. 

# Dataset

Criteria for an appropriate journal article are:

1. Data are publicly available
2. The authors' code (R or otherwise) is **not** available.
3. Authors used regression for analysis.
4. You have not used these data before.
5. It is a relatively complicated dataset. Ways that a dataset can be "relatively complicated" are:
    a. There are at $\gtrapprox$ 20 predictor variables, or
    b. The authors use a method not taught in this class (e.g. principal components regression, lasso, etc).
  
I will let you propose a journal article with a dataset, but I retain the right to decide if the dataset is "relatively complicated".

[PLoS](https://journals.plos.org) Journals are all required to have a data availability statement, so this is a good place to start searching.

You can get a sense of the type of articles I am looking for (or use one of these) based on the following that I picked out of a [search from PLoS One](https://journals.plos.org/plosone/search?filterJournals=PLoSONE&q=%22linear%20regression%22&sortOrder=DATE_NEWEST_FIRST&page=1&utm_content=a&utm_campaign=ENG-467). These aren't necessarily good or interesting articles, just ones that satisfy the above requirements and were released in July 2022 when I checked for the newest releases in PLoS One.

If the data are in a weird format (like "sav" or "dta"), then try to use the [`{haven}`](https://cran.r-project.org/package=haven) package to load it into R.

- @goli2022estimates: 
    - This project seems to require a lot of data download and data munging to obtain the same variables as in the study.
    - You don't need to do the spatial stuff in this paper (a little too advanced). But let me know if you want to try!
- @eseta2022factors
    - They perform principal components analysis (PCA) to extract features.
    - The data are sav.
- @yin2022does
    - They use factor analysis (related to PCA).
    - No need to do the path analysis.
- @haile2022predictors
    - The data are dta.
- @taillie2022developing
    - One of their analyses uses mixed-effects modeling.
    - The data are dta.
- @karnaze2022association
    - They do both logistic and linear regressions here.
    - The data are sav.
- @tekle2022length
    - The data are sav.
- @hettich2022conspiracy
- @lu2022emojis
    - Only do this one if you have had some machine learning experience.

# Progress Report (5 pts)

- You should turn in a 1 page written report affirming:

    1. That you have downloaded the data and have successfully loaded it into R.
    2. That you have checked that all variables described in the paper are available in the dataset.
    3. That you have made scatterplots and histograms of the variables in the dataset.
    4. If a new method is needed (e.g. PCA), that you have downloaded a tutorial from the web and can now implement the new method in R code.
    
- I want to see evidence that all of the above has been accomplished.

# Presentation/Written Report

You will provide a presentation and a written report. They should both follow this outline:

- Introduce the problem/paper.
- Describe the data.
- State the model(s)/method(s)
- Check assumptions.
- Provide interpretation of results.
- Critiques / expansion of analyses.
- Summary/Conclusion

## Presentation (10 pts)

- Your presentation should be 15 minutes long.

- All group members should speak.

- I will ask questions at the end of the presentation.

## Final Written Report (10 pts)

- The text of your final written report should be max 4 pages, double spaced, 12 point font. This does not include figures/tables/references/code.

# Code (5 pts)

- You should turn in all of your code to reproduce the results of the paper in a single zipped folder. 

- I should be able to run your code without modification and obtain your results.

- There should be no errors in your code. 

- You should use the methods/packages we've learned in class.

# Peer Assessment (2 pts)

At the end of the project, I will have your teammates rate your contribution to the project. I will adjust your grade according to their comments. If they all say that you didn't do anything, I will give you a failing grade.

You will get 2 points for filling out the assessment.

# References
