---
title: "Final Project"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
bibliography: "source.bib"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Important Dates

# General Description

- For the final project you will work in teams of 2--3 to select a journal article from 2022 and reproduce their results. 

# Dataset

Criteria for an appropriate journal article are:

1. Data are publicly available
2. The authors' code (R or otherwise) is **not** available.
3. Authors used regression for analysis.
4. You have not used these data before.
5. It is a relatively complicated dataset. Ways that a dataset can be "relatively complicated" are:
    a. There are at least 20 predictor variables, or
    b. The authors use a method not taught in this class (e.g. principal components regression, lasso, etc).
  
I will let you propose a journal article with a dataset, but I retain the right to choose if the dataset is "relatively complicated".

[PLOS](https://journals.plos.org) Journals are all required to have a data availability statement, so this is a good place to start searching.

You can get a sense of the type of articles I am looking for (or use one of these) based on the following that I picked out of a [search from PLOS](https://journals.plos.org/plosone/search?filterJournals=PLoSONE&q=%22linear%20regression%22&sortOrder=DATE_NEWEST_FIRST&page=1&utm_content=a&utm_campaign=ENG-467).

- @karat2022estimating
- @goli2022estimates
- @dey2022prediction
- @eseta2022factors
- @yin2022does
- @haile2022predictors
- @taillie2022developing

# Progress Report

# Presentation

# Final Written Report

- Describe analysis.
- State model
- Check assumptions
- Provide interpretation
- Critiques (expand on analysis if so).

# Grade Breakdown

# References
