---
title: "Final Project"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
bibliography: "source.bib"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Important Dates

- 04/11/2024: Journal article selected.
- 04/25/2024: Progress report due.
- 05/07/2024 (Section 001) or 05/08/2024 (Section 002): 
    - Final project presentations. 
    - Written report due.
    - Code due.
    - All of this is due **before** the final project presentations (**not** on the following Thursday).

# General Description

- For the final project you will work in teams of 2--3 to select a journal article from 2023 or later and reproduce their results. 

# Dataset

Criteria for an appropriate journal article are:

1. Data are publicly available
2. The authors' code (R or otherwise) is **not** available.
3. Authors used regression for analysis.
4. You have not used these data before, and are not using them for anything else (including other classes) this semester.
5. The data are from 2023 or later.
6. No other group is using that data.
7. It is a relatively complicated dataset. Ways that a dataset can be "relatively complicated" are:
    a. There are at $\gtrapprox$ 10 predictor variables, or
    b. The authors use a method not taught in this class (e.g. principal components regression, lasso, etc).
  
I will let you propose a journal article with a dataset, but I retain the right to decide if the dataset is "relatively complicated".

[PLoS](https://journals.plos.org) Journals are all required to have a data availability statement, so this is a good place to start searching.

You can get a sense of the type of articles I am looking for (or use one of these) based on the following that I picked out of a [search from PLoS One](https://journals.plos.org/plosone/search?filterJournals=PLoSONE&q=%22linear%20regression%22&sortOrder=DATE_NEWEST_FIRST&page=1&utm_content=a&utm_campaign=ENG-467). These aren't necessarily good or interesting articles, just ones that satisfy the above requirements and were released when I checked for the newest releases in PLoS One.

If the data are in a weird format (like "sav" or "dta"), then try to use the [`{haven}`](https://cran.r-project.org/package=haven) package to load it into R.


- @fu2024unveiling
- @zhao2024sex
- @frikha2024towards
- @khanthik2024factors
- @shinozuka2024peripapillary
- @wei2024influence
- @ridder2024art
- @xuan2024nonlinear
- @vejrik2024non
- @villiers2024sleep


# Progress Report (10 pts)

- It often (not always) takes about two weeks to reproduce the results of a paper. 

- Sometimes, you find out right away that reproducing the paper is impossible, and you need to switch papers. This takes up time.

- For the progress report you will turn in an R markdown file that loads in the data (that is already cleaned by some other script), fits a regression and reproduces the tables in the paper you have selected.

- If your numbers are not the same as in the paper, your options are:
    1. Switch papers.
    2. Convince me that both (i) the results are impossible to reproduce with the given data and (ii) you can still get something interesting out of the data that you do have. If I'm not convinced of both of these, you'll have to switch papers.

- From past experience, students don't believe me when I say it often takes two weeks to reproduce the results of a paper, so they start the day before the progress report is due and ask for extensions.

- **There will be no extensions** on the progress report, even if you need to switch papers. Instead, there will be an upper bound on the grade of your progress report grade based on how late you turn it in.

- < Two days late: 9 out of 10 pts max
- < Four days late: 8 out of 10 pts max
- < Six days late: 7 out of 10 pts max
- < Eight days late: 6 out of 10 pts max
- < Ten days late: 5 out of 10 pts max
- More than ten days late: 0 out of 10 pts

# Presentation/Written Report

You will provide a presentation and a written report. They should both follow this outline:

- Introduce the problem/paper.
- Describe the data.
- State the model(s)/method(s)
- Check assumptions.
- Provide interpretation of results.
- Critiques / expansion of analyses.
- Summary/Conclusion

## Presentation (5 pts)

- Your presentation should be 15 minutes long.

- All group members should speak.

- I will ask questions at the end of the presentation.

## Final Written Report (5 pts)

- This should be an executive summary with lots of bullet points. Don't make me read text not immediate relevant to the project. No yapping.
    - Summarize paper's goals.
    - Summarize data.
    - Summarize model.
    - Check assumptions.
    - Interpret results.
    - Provide critiques (this is what I am most interested in)
    - Summarize your findings.

- Limit your write-up to 4 pages, double spaced, 12 point font. This does not include figures/tables/references/code.

# Code (5 pts)

- You should turn in all of your code to reproduce the results of the paper in a single zipped folder. 

- I should be able to run your code without modification and obtain your results.

- There should be no errors in your code. 

- You should use the methods/packages we've learned in class.

# Peer Assessment (2 pts)

At the end of the project, I will have your teammates rate your contribution to the project. I will adjust your grade according to their comments. If they all say that you didn't do anything, I will give you a failing grade.

You will get 2 points for filling out the assessment.

# References
