---
title: "R Graphics with `{ggplot2}`"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 4, 
                      fig.height = 3, 
                      fig.align = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

```{r, echo=FALSE}
set.seed(1)
```

# Learning objectives

- Basic plotting in R using the `{ggplot2}` package.

# Introduction

- `{ggplot2}` is very powerful, so I am just going to show you the most important and basic plots that are necessary for data analysis.

- Before using the plotting functions from `{ggplot2}` in a new R session, always first load the `{ggplot2}` library.

    ```{r}
    library(ggplot2)
    ```

- In this vignette, we'll also make some variable transformations, so we will need the `{dplyr}` package.

    ```{r, message=FALSE}
    library(dplyr)
    ```

- I will use the `mpg` dataset to demonstrate plots

    ```{r}
    data(mpg, package = "ggplot2")
    glimpse(mpg)
    ```

# `ggplot()`

- The first function you use in making a plot is always `ggplot()`.

- It takes two main arguments:
    - `data`: The data frame that holds the variables you want to plot.
    - `mapping`: The "aesthetic map"
    
- An "aesthetic map" says what variables go on the $x$-axis, what variables go on the $y$-axis, what variables are represented by color, or size, or point shape, etc...

- You place all aesthetic maps inside an `aes()` function.

- E.g. here, we are mapping `hwy` to be on the $x$-axis, and different values of `drv` should be different colors.

    ```{r, eval = FALSE}
    ggplot(data = mpg, mapping = aes(x = hwy, color = drv))
    ```
    
- This function just sets the data and the aesthetic mapping, but it won't produce any useful plot by itself.

- You add additional functions to the plot to state the type of plot you want.

# One Variable

## Continuous

- Histogram:
    - Variable should be on the $x$-axis.
    - Use the `geom_histogram()` function.

    ```{r, message=FALSE}
    ggplot(data = mpg, mapping = aes(x = hwy)) +
      geom_histogram()
    ```

- Make the bin lines black and the fill white, and change the number of bins.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = hwy)) +
      geom_histogram(bins = 10, color = "black", fill = "white")
    ```

- **Exercise**: Load in the [estate](https://dcgerard.github.io/stat_415_615/data/estate.csv) data (see [here](https://dcgerard.github.io/stat_415_615/data.html#real-estate-sales) for a description) and make a histogram of price with 20 bins. Make the bins red.

    ```{r, echo = FALSE, eval = FALSE}
    library(readr)
    library(ggplot2)
    estate <- read_csv("https://dcgerard.github.io/stat_415_615/data/estate.csv")
    ggplot(data = estate, mapping = aes(x = price)) +
      geom_histogram(bins = 20, fill = "red")
    ```

## Discrete

- Barplot:
    - Put the variable on the $x$-axis.
    - Use `geom_bar()`.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = drv)) +
      geom_bar()
    ```

- **Exercise**: What variables from the `estate` data are appropriately plotted using a bar plot? Plot them.

    ```{r, eval = FALSE, echo = FALSE}
    ggplot(data = estate, mapping = aes(x = quality)) +
      geom_bar()
    ggplot(data = estate, mapping = aes(x = style)) +
      geom_bar()
    ggplot(data = estate, mapping = aes(x = ac)) +
      geom_bar()
    ggplot(data = estate, mapping = aes(x = pool)) +
      geom_bar()
    ggplot(data = estate, mapping = aes(x = highway)) +
      geom_bar()
    ```

# Two Variables

## Continuous X, Continuous Y

- Scatterplot:
    - Say what variables should be on the $x$- and $y$-axes. 
    - Use `geom_point()`.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point()
    ```

- Jitter points to account for overlaying points.
    - Use `geom_jitter()` instead of `geom_point()`.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_jitter()
    ```

- Add a Loess Smoother by adding `geom_smooth()`.

    ```{r, message=FALSE}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() +
      geom_smooth(se = FALSE)
    ```

- **Exercise**: Using the `estate` data, make a scatterplot of number of bedrooms versus number of bathrooms. Adjust for any overplotting.

    ```{r, echo = FALSE, eval = FALSE}
    ggplot(data = estate, mapping = aes(x = bed, y = bath)) +
      geom_jitter()
    ```

## Discrete X, Continuous Y

- Boxplot
    - Place one variable on $x$-axis and other on $y$-axis.
    - Typically, but not always, continuous goes on $y$-axis.
    - Use `geom_boxplot()`.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
      geom_boxplot()
    ```

- **Exercise**: Using the `estate` data, plot sales price versus style. (*hint*: you need to first convert `style` to a factor using `as.factor()`)

    ```{r, eval = FALSE, echo = FALSE}
    estate <- mutate(estate, style = as.factor(style))
    ggplot(data = estate, mapping = aes(x = style, y = price)) +
      geom_boxplot()
    ```

# Color Coding and Legends

- Color code a scatterplot by a categorical variable and add a legend.
    - Just add a color mapping.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = drv)) +
      geom_jitter()
    ```

- **Exercise**: Using the `estate` data, create a boxplot of price versus ac, color coding by pool.

    ```{r, echo = FALSE, eval = FALSE}
    ggplot(data = estate, mapping = aes(x = ac, y = price, color = pool)) +
      geom_boxplot()
    ```

# Changing a legend title

- Add a `scale_*()` call to change the name:

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy, color = drv)) +
      geom_jitter() +
      scale_color_discrete(name = "New Name1")
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy, shape = drv)) +
      geom_jitter() +
      scale_shape_discrete(name = "New Name2")
    ```

# Faceting

- You can facet by a categorical variable by adding a `facet_grid()` or `facet_wrap()` function.

- The variable to the left of the tilde ("`~`") indexes the row facets, the variable to the right of the tilde indexes the column facets. Using a dot ("`.`") in place of a variable means that there will only be one row/column facet.

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() +
      facet_grid(. ~ drv)
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() +
      facet_grid(drv ~ .)
    ```
    
    ```{r, fig.height=2.7, fig.width=3.5}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() +
      facet_grid(fl ~ drv)
    ```

- **Exercise**: Using the `estate` data, plot price versus area, faceting by ac, color coding by pool.

    ```{r, echo = FALSE, eval = FALSE}
    ggplot(data = estate, mapping = aes(x = area, y = price, color = pool)) +
      geom_point() +
      facet_wrap(. ~ ac)
    ```

# Change Theme

- Add a `theme_*()` function to change the theme:

    ```{r}
    ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
      geom_point() + 
      theme_classic()
    ```
