---
title: "R Graphics with `qplot`"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 4, 
                      fig.height = 3, 
                      fig.align = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

```{r, echo=FALSE}
set.seed(1)
```

# Learning objectives

- Basic plotting in R using `qplot()` from the `{ggplot2}` package.

# Introduction

- `{ggplot2}` is very powerful, so I am just going to show you the most important and basic plots that are necessary for data analysis.

- Before using `qplot()` in a new R session, always first load the `{ggplot2}` library.

    ```{r}
    library(ggplot2)
    ```

- In this vignette, we'll also make some variable transformations, so we will need the `{dplyr}` package.

    ```{r, message=FALSE}
    library(dplyr)
    ```

- I will use the `mpg` dataset to demonstrate plots

    ```{r}
    data(mpg, package = "ggplot2")
    glimpse(mpg)
    ```

# `qplot()`

- All plots will be created with `qplot()`. 

- You specify the type of plot with the `geom` argument (for "geometry").

- You specify which variables map to features of this plot via the other arguments.

- E.g. to say that `hwy` maps to the x-axis, use `x = hwy`. To say that it maps to the color, use `color = hwy`, etc.

- To map a constant to a feature (e.g. "red" to the color), wrap "red" in `I()`. E.g. `color = I("red")`.

# One Variable

## Continuous

- Histogram

    ```{r}
    qplot(x = hwy, data = mpg, geom = "histogram", bins = 10)
    ```

- Make the bin lines black and the fill white.

    ```{r}
    qplot(x = hwy,
          data = mpg,
          geom = "histogram",
          bins = 10, 
          color = I("black"),
          fill = I("white"))
    ```

- **Exercise**: Load in the [estate](https://dcgerard.github.io/stat_415_615/data/estate.csv) data (see [here](https://dcgerard.github.io/stat_415_615/data.html#real-estate-sales) for a description) and make a histogram of price with 20 bins. Make the bins red.

    ```{r, echo = FALSE, eval = FALSE}
    library(readr)
    library(ggplot2)
    estate <- read_csv("https://dcgerard.github.io/stat_415_615/data/estate.csv")
    qplot(x = price, data = estate, bins = 20, fill = I("red"))
    ```

## Discrete

- Barplot

    ```{r}
    qplot(x = drv, data = mpg, geom = "bar")
    ```

- **Exercise**: What variables from the `estate` data are appropriately plotted using a bar plot? Plot them.

    ```{r, eval = FALSE, echo = FALSE}
    qplot(x = quality, data = estate, geom = "bar")
    qplot(x = style, data = estate, geom = "bar")
    qplot(x = ac, data = estate, geom = "bar")
    qplot(x = pool, data = estate, geom = "bar")
    qplot(x = highway, data = estate, geom = "bar")
    ```

# Two Variables

## Continuous X, Continuous Y

- Scatterplot

    ```{r}
    qplot(x = cty, y = hwy, data = mpg, geom = "point")
    ```

- Jitter points to account for overlaying points.

    ```{r}
    qplot(x = cty, y = hwy, data = mpg, geom = "jitter")
    ```

- Add a Loess Smoother

    ```{r, message=FALSE}
    qplot(x = cty, y = hwy, data = mpg, geom = "jitter") + 
      geom_smooth()
    ```

- **Exercise**: Using the `estate` data, make a scatterplot of number of bedrooms versus number of bathrooms. Adjust for any overplotting.

    ```{r, echo = FALSE, eval = FALSE}
    qplot(x = bed, y = bath, data = estate, geom = "jitter")
    ```


## Discrete X, Continuous Y

- Boxplot

    ```{r}
    qplot(x = class, y = hwy, data = mpg, geom = "boxplot")
    ```

- **Exercise**: Using the `estate` data, plot sales price versus style. (*hint*: you need to first convert `style` to a factor using `as.factor()`)

    ```{r, eval = FALSE, echo = FALSE}
    estate <- mutate(estate, style = as.factor(style))
    qplot(x = style, y = price, data = estate, geom = "boxplot")
    ```

# Color Coding and Legends

- Color code a scatterplot by a categorical variable and add a legend.

    ```{r}
    qplot(x = cty, y = hwy, color = drv, data = mpg, geom = "jitter")
    ```

- **Exercise**: Using the `estate` data, create a boxplot of price versus ac, color coding by pool.

    ```{r, echo = FALSE, eval = FALSE}
    qplot(x = ac, y = price, color = pool, data = estate, geom = "boxplot")
    ```

# Changing a legend title

- Add a `scale_*()` call to change the name:

    ```{r}
    qplot(x = cty, y = hwy, color = drv, data = mpg, geom = "jitter") + 
      scale_color_discrete(name = "New Name1")
    qplot(x = cty, y = hwy, shape = drv, data = mpg, geom = "jitter") +
      scale_shape_discrete(name = "New Name2")
    ```

# Faceting

- You can facet by a categorical variable using the `facets` argument.

- The variable to the left of the tilde ("`~`") indexes the row facets, the variable to the right of the tilde indexes the column facets. Using a dot ("`.`") in place of a variable means that there will only be one row/column facet.

    ```{r}
    qplot(cty, hwy, data = mpg, facets = . ~ drv, geom = "point")
    qplot(cty, hwy, data = mpg, facets = drv ~ ., geom = "point")
    ```
    
    ```{r, fig.height=2.7, fig.width=3.5}
    qplot(cty, hwy, data = mpg, facets = fl ~ drv, geom = "point")
    ```

- **Exercise**: Using the `estate` data, plot price versus area, faceting by ac, color coding by pool.

    ```{r, echo = FALSE, eval = FALSE}
    qplot(x = area, y = price, color = pool, facets = .~ac, geom = "point", data = estate)
    ```

# Change Theme

- Add a `theme_*()` function to change the theme:

    ```{r}
    qplot(x = cty, y = hwy, data = mpg) +
      theme_bw()
    ```

# General Considerations

The main options that I use are

- Options for `"geom"` argument:
    - `"point"`: Makes scatterplots.
    - `"line"`: Makes a line plot.
    - `"histogram"`: Makes a histogram.
    - `"boxplot"`: Makes a boxplot.
    - `"density"`: Makes the density plot.
    - `"bar"`: First tabulates frequencies of each value, then makes a barplot.
    - `"smooth"`: Fits a smooth line to a cloud of points and plots the output.
    - `"dotplot"`: Makes a dotplot.

`qplot` has other arguments that control the way the plot looks. You should read about these arguments. In particular, read carefully the help page `?qplot`. Useful ones are:

- `data`: Specify the dataframe that all variables belong to.
- `main`: This controls the title.
- `xlab`, `ylab`: These control the x and y axis labels.
- `color`: Controls the color of the lines/points.
- `fill`: Controls the color of areas (e.g. for histograms).
- `size`: Controls the size of points.
- `shape`: The shape of points (`"circle"`, `"square"`, `"triangle"`, etc...)
- `alpha`: Controls the level of transparency of points/lines/fills.
- `lwd`: Line width.
- `lty`: Line type (`"solid"`, `"dashed"`, `"dotted"`, etc...).
- `facets`: Split up the data into multiple plots.

If you want to make all points the same shape/size/color, you need to enclose the size/shape/color using the function `I()`.

If a variable is being treated as continuous rather than categorical, you need to enclose that variable in a `factor()` function call.
