---
title: "Final Project"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Synopsis

You will work in groups of 3 or 4 people on a real data analysis project. Each group member is expected to be familiar with every aspect of the project, even the parts that they did not work on. There will be a within-group evaluation at the end of the project that will count for a portion of the project grade.

You will build multiple regression models to analyze a dataset from the real world to address a research question of interest. You are responsible for choosing this topic. I will not accept datasets from a publication, or that have already been deeply analyzed by others (like most Kaggle datasets). Here are some possible sources:

- [College Scorecard](https://collegescorecard.ed.gov/data/)
- [USA Spending](https://www.usaspending.gov/)
- [World Bank](https://data.worldbank.org/)

You will submit the following on Canvas. 

a. Project proposal (1--2 pages). It should cover, but is not limited to, the question of interest, the variables under consideration, the methods to collect data, proposed models, etc.
b. Progress report (1--2 pages). It should cover, but is not limited to, the work you’ve done so far and your next steps.
c. The data set for your analysis, the articles being reviewed and a list of references. 
d. The presentation material (power point or notes/handouts).
e. A written report (2--3 pages). It should cover, but is not limited to, your question(s) of interest, your final regression model(s), verification of model(s) assumptions, and conclusions.

# Extra Project for 615

If you are taking this course for 615 credit, then I want to you also present on a topic from linear regression that we have not covered extensively. No written report is needed here, you will just prepare a 20 minute lecture to give to the class. Work in groups of 3 or 4. Possible topics include:

- LASSO
- Random Effects Models
- Genome-wide Association Studies
- Regression splines
- Quantile regression

You are responsible for choosing the topic of the project and looking for appropriate reference material. I will serve as your "consultant" throughout your project.

At the end of the semester, each group will give a 20 minutes' lecture to introduce the method you have studied. You can give the lecture in your own style. However, it should include, but is not limited to, the following:

a. An introduction of the statistical model, including the assumptions.
b. A discussion on when and where the method can be applied. Give an example.
c. The interpretation of analysis results from the method.
d. A discussion on the benefits and cautions of the method.
e. Extra credit: example and/or software demonstration.

