---
title: "Final Project"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Important Dates

- 10/07/2021: Project Proposal Due
- 11/11/2021: Progress Report Due.
- 12/14/2021 or 12/15/2021: Group project presentations, and final reports due.

# General Description

- Everyone will work in groups of 3 or 4 people to build multiple regression models to analyze a dataset from the real world to address a research question of interest, culminating in a written report and a project presentation.

- Graduate students (those taking credit for 615), should also prepare a group presentation on a topic from regression analysis that we did not cover in class. They may work with the same or different people as they worked with in the analysis project.

# Project Proposal

- You are responsible for choosing the topic of analysis. I will not accept datasets from a publication, or those that have already been deeply analyzed by others (like most Kaggle datasets). Here are some possible sources:

    - [College Scorecard](https://collegescorecard.ed.gov/data/)
    - [USA Spending](https://www.usaspending.gov/)
    - [US Census Data](https://data.census.gov/)
    - [World Bank](https://data.worldbank.org/)

- For the project proposal, you will submit to Canvas:
    1. A 1--2 page proposal that should cover, but is not limited to
        a. The question of interest,
        b. The data to be analyzed,
        c. The variables under consideration.
    2. The dataset you will analyze in CSV format.

# Progress Report

- Midway through the semester, you will submit a progress report.

- This should be a 1--2 page report detailing the results of your exploratory data analysis. This should include:
    1. What variables you consider to be important for the problem at hand.
    2. Some plots demonstrating that you are informally exploring associations.
    3. A description of your next steps.
    
# Presentation

- Your group will prepare a 15 minute presentation for the class.


- When preparing your presentation, consider these guidelines:
    1. Your presentation should focus on solving the applied problem, and it should be accessible to the 
    "educated layperson" audience. 
    2. You should concisely and clearly describe the research question of interest and its background.  
    3. You should explain the design of the study and how the data are collected. This would include, but is not limited to, a discussion of study design, sampling process, target population, etc., where applicable. 
    4. You should concisely and clearly describe and present the data. Use appropriate exploratory statistics and graphs to summarize the data.  
    5. You should present the statistical methods used for analysis, including a brief discussion of advantages, disadvantages and necessary assumptions of methods.  
    6. You should make statistical inferences on the data and discuss the statistical results in the context of the problem.  
    7. You should make concise conclusions in the context of the problem and discuss limitations of the analysis, if applicable.  

# Final Written Report

- You should prepare a written report (2--3 pages) that should cover (but is not limited to)
    1. Your question(s) of interest, 
    2. Your final regression model(s), 
    3. Verification of model(s) assumptions, and
    4. conclusions.

- Use the same guidelines as above in the presentation section when preparing your final written report.

# Extra Project for 615

If you are taking this course for 615 credit, then I want to you also present on a topic from linear regression that we have not covered extensively. No written report is needed here, you will just prepare a 10 minute lecture to give to the class. Work in groups of 3 or 4. Possible topics include:

- [LASSO](https://en.wikipedia.org/wiki/Lasso_(statistics))
- [Random Effects Models](https://en.wikipedia.org/wiki/Random_effects_model)
- [Genome-wide Association Studies](https://en.wikipedia.org/wiki/Genome-wide_association_study)
- [Quantile regression](https://en.wikipedia.org/wiki/Quantile_regression)
- [Principal Components Analysis](https://en.wikipedia.org/wiki/Principal_component_analysis)
- [Bayesian Linear Regression](https://en.wikipedia.org/wiki/Bayesian_linear_regression)

You are responsible for choosing the topic of the project and looking for appropriate reference material. I will serve as your "consultant" throughout your project.

At the end of the semester, each group will give a 10 minute lecture to introduce the method you have studied. You can give the lecture in your own style. However, it should include, but is not limited to, the following:

a. An introduction of the statistical model, including the assumptions.
b. A discussion on when and where the method can be applied. Give an example.
c. The interpretation of analysis results from the method.
d. A discussion on the benefits and cautions of the method.
e. (BONUS) Software demonstration.

# Grade Breakdown

50 pts total

- 415 Undergraduate Students
    - (10 pts) Project proposal, team and data selection.
    - (10 pts) Progress report.
    - (10 pts) Final report
    - (20 pts) Final presentation
- 615 Graduate Students
    - (10 pts) Project proposal, team and data selection.
    - (10 pts) Progress report.
    - (10 pts) Final report
    - (10 pts) Analysis Final presentation
    - (10 pts) Method Final presentation
