---
title: "Equation Review"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
urlcolor: "blue"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Summary

Here is a list of the equations I want you to have memorized at the top of your head.

# List

- Mean: $\bar{Y} = \frac{1}{n}\sum_{i=1}^n$
- SLR Model: $Y_i = \beta_0 + \beta_1 X_i + \epsilon_i$, $E[\epsilon_i] = 0$, $var(\epsilon_i) = \sigma^2$, $cor(\epsilon_i, \epsilon_j) = 0$ for $i\neq j$.
- Normal SLR Model: $Y_i = \beta_0 + \beta_1 X_i + \epsilon_i$, $\epsilon \overset{iid}{\sim} N(0, \sigma^2)$.
- OLS Objective: $\sum_{i=1}^n\left[Y_i - (\beta_0 + \beta_1X_i)\right]^2
- SLR OLS estimates: $\hat{\beta}_0 = \bar{Y} - \hat{\beta}_1\bar{X}$, $\hat{\beta}_1 = cor(X, Y)\frac{sd(Y)}{sd(X)}$.
- Residuals $e_i = Y_i - \hat{Y}_i$.
- Properties of fitted regression line: (i) mean of residuals is 0, (ii) mean of observed values equals mean of fitted values, (iii) residuals are uncorrelated with predictors, (iv) residuals are uncorrelated with *fitted* values, and (v) regression line always goes through mean $(\bar{X}, \bar{Y})$.
- MSE: $MSE = \frac{1}{n-p}\sum_{i=1}^n\left[Y_i - \hat{Y}_i\right]^2$.
- $t$-statistic: $t^* = \frac{\hat{\beta}_1}{s(\hat{\beta}_1)}$, which follows a $t_{n-p}$ distribution numder the null that $\beta_1 = 0$.
- Get two-sided $p$-value manually via $2 * pt(q = -abs(t^*), df = n - p)$.
- Confidence interval = estimate $\pm$ multiplier $\times$ standard error
    - Typically, multiplier = $qt(1 - \alpha/2, df)$, tell me what $\alpha$ and $df = n-p$ should be.
    - In prediction interval, $s^2(pred) = s^2(\hat{Y}_i) + MSE$.
- $SSE = \sum_{i=1}^n(Y_i - \hat{Y}_i)^2$, with $df_E = n - p$
- $SSR = \sum_{i=1}^n(\hat{Y}_i - \bar{Y}_i)^2$, with $df_R = n - p$ 
- $SSTO = \sum_{i=1}^n(Y_i - \bar{Y}_i)^2$, with $df_{TO} = n-1$
- $SSTO = SSE + SSR$
- $F^* = \frac{SSR/df_R}{SSE/df_E}$, which follows an $F(df_R, df_E)$ distribution under the null model of $Y_i = \beta_0 + \epsilon_i$.
- $R^2 = \frac{SSR}{SSTO} = 1 - \frac{SSE}{SSTO}$

