---
title: "MLR III: Practical Considerations"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
library(latex2exp)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Learning Objectives

- Concretely describe the proper steps of any analysis.
- Choice of scaling predictors for interpretation.
- Effect sizes, practical significance, and statistical significance.
- Chapter 12 from ROS

# Steps of an analysis

- For **any analysis** where you are applying a model, the steps you should take are:
    1. Exploratory data analysis
    2. Choose the form of the model.
    3. Fit the model to the data.
    4. Assess how well the model describes the data.
    5. Use the model to address the question of interest.
    
- You should be performing each of these steps when you apply a model to data.

# Choosing a scale for interpretation

- You should apply linear scalings to your data.

- Height in millimeter (difference in $Y$ for a 1 mm difference in height?) versus height in centimeters (difference in $Y$ for a 1 cm difference in height?) versus height in meters (difference in $Y$ for a 1 m difference in height?). For humans, cm is probably the best scale.

- Use standard deviations to guide your choice of scale. Roughly, 68\% of observations will be within 1 standard deviation of the mean, and so choosing your predictor on the order standard deviations (so within a magnitude of the SD) makes coefficients more interpretable.

# Practical versus statistical significance

- Tests can have tiny $p$-values, but the effect sizes might be small.

# Comparing effect sizes to residual standard error

- It is always a good idea to compare the effect sizes to the residual standard error.

- **Example**: Recall the [earnings data](https://dcgerard.github.io/stat_415_615/data.html#Earnings_Data). Let's read it in and fit a model for log-earnings on height:

    ```{r}
    library(tidyverse)
    earnings <- read_csv("https://dcgerard.github.io/stat_415_615/data/earnings.csv")
    earnings <- mutate(earnings, log_earn = log(earn))
    earnings <- filter(earnings, is.finite(log_earn))
    qplot(x = height, y = log_earn, data = earnings)
    lm_earn <- lm(log_earn ~ height, data = earnings)
    ```
    
    ```{r}
    library(broom)
    tidy(lm_earn)
    ```
    
    The $p$-value is **tiny**. So is this a huge effect? Not really. Let's discuss.
    
    Individuals that are a whole foot taller earn about $0.05704 \times 12 = 0.6845$ log-dollars more. This corresponds to about twice as much money ($e^0.6845 = 1.98). This seems large. However, let's look at the residual standard deviation.
    
    ```{r}
    glance(lm_earn)
    ```
    
    The residual SD is 0.8772. This is the average variability about the regression line and is larger even when comparing folks that are a full foot different in height.
    
    If we compare prediction intervals between a 5'2'' individual and a 6'2'' individual (ignoring the appropriateness of the normal model for now), we have
    
    ```{r}
    predict(object = lm_earn, 
            newdata = data.frame(height = c(5.2 * 12, 6.2 * 12)),
            interval = "prediction") %>%
      exp()
    ```
    So it is true that the taller individual is expected to make twice as much money (\$12,938 versus \$25,652), but the range of individuals at each level is huge (\$2312 to \$72,387 and \$4579 to \$143,718). And this is on the larger side of comparisons between individuals (most individuals are less than a foot different).
    



