---
title: "Linear Algebra"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
library(latex2exp)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center")
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Learning Objectives

- Multiple linear regression model definition and interpretation.
- Indicator variables, transformed variables, interaction effects,
- Inference
- Diagnostics and remedial measures.
- Chapter 6 of KNNL

# Overview

- The multiple linear regression (MLR) model is of the form

    \begin{align}
    Y_i &= \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + \cdots + \beta_pX_{ip} + \epsilon_i\\
    E[\epsilon_i] &= 0\\
    var(\epsilon_i) &= \sigma^2\\
    cov(\epsilon_i, \epsilon_j) &= 0 \text{ for all } i \neq j
    \end{align}

- This lecture extends everything we know from SLR To MLR.

- The next few lectures cover issues specific to MLR that do not appear in SLR.
