---
title: "ANOVA View of Linear Models"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
library(latex2exp)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Learning Objectives

- ANOVA Perspective of hypothesis testing in the multiple linear regression model.
- Sections 6.5, 6.9, 7.1, 7.2, and 7.3 of KNNL

# $F$-test for lack of fit

- If you have repeat observations at the same levels of predictor variables, then you can run a lack-of-fit test as before.

- NOTE: You need to have repeat observations where *all* predictor levels are equal. E.g. this design matrix has *no* repeat observations:
    $$
    \mathbf{X} = 
    \begin{pmatrix}
    1 & 55 & 13\\
    1 & 55 & 17\\
    1 & 43 & 13
    \end{pmatrix}
    $$
    whereas this design matrix has repeat observations at one level (55 for the first predictor and 13 at the second).
    $$
    \mathbf{X} = 
    \begin{pmatrix}
    1 & 55 & 13\\
    1 & 55 & 13\\
    1 & 43 & 13
    \end{pmatrix}
    $$
