---
title: "SLR Case Study"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
urlcolor: "blue"
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Learning Objectives

- Provide an example of how I would approach a simple linear regression task.

# Data Analysis

- We will use the Plutonium data from Section 3.11 of the book:
    ```{r, message = FALSE}
    library(tidyverse)
    library(broom)
    plut <- tribble(~act, ~alpha,
                    0.150, 20,
                    0.004, 0,
                    0.069, 10,
                    0.030, 5,
                    0.011, 0,
                    0.004, 0,
                    0.041, 5,
                    0.109, 20,
                    0.068, 10,
                    0.009, 0,
                    0.009, 0,
                    0.048, 10,
                    0.006, 0,
                    0.083, 20,
                    0.037, 5,
                    0.039, 5,
                    0.132, 20,
                    0.004, 0,
                    0.006, 0,
                    0.059, 10,
                    0.051, 10,
                    0.002, 0,
                    0.049, 5,
                    0.106, 0
                    )
    ```
    
- The two variables are
    - `act`: A measure of the plutonium activity, measured in [picocuries](https://en.wikipedia.org/wiki/Curie_(unit)) (one trillianth of a curie) per gram.
    - `alpha`: The intensity of alpha particle strikes in counts per second.
    
- **The first thing you should always do is plot your data**.

    ```{r}
    qplot(x = alpha, y = act, data = plut)
    ```

- There is one clear outlier. If I was collaborating with someone who collected the data, I would ask them about that point. The authors did so and it was determined that lab conditions were not properly maintained for that observation, so we'll remove it.

    ```{r}
    plut <- filter(plut, !(alpha == 0  & act > 0.1))
    qplot(x = alpha, y = act, data = plut)
    ```
    
    
- Let's use a smoother to explore the relationship.

    ```{r, message = FALSE, warning = FALSE}
    plut <- filter(plut, !(alpha == 0  & act > 0.1))
    qplot(x = alpha, y = act, data = plut) +
      geom_smooth(se = FALSE)
    ```

- The relationship looks fairly linear. Let's try a linear regression and explore the residuals.

    ```{r}
    lmout <- lm(act ~ alpha, data = plut)
    aout <- augment(lmout)
    qplot(x = .fitted, y = .resid, data = aout) +
      geom_hline(yintercept = 0)
    ```

- We have a non-constant variance issue. What I would do now would depend on the goal of the research. If my collaborators were interested in just the relationship between alpha counts and activity level, then I would use sandwich estimation of the standard errors.

    ```{r, message = FALSE}
    library(sandwich)
    library(lmtest)
    cout <- coeftest(x = lmout, vcov. = vcovHC(x = lmout))
    tidy(cout, conf.int = TRUE)
    ```
    
    I would conclude that we have strong evidence of a linear association between alpha count and plutonium activity ($p < 0.001$, $n = 23$). Readings of 10 more counts are estimated to have activities about 0.055 pCi/g higher (95\% CI of 0.041 pCi/g to 0.070 pCi/g higher).

- However, if the goal is to provide prediction, along with interval estimates of those predictions, then I would try working a little harder. First, a log-transformation of $y$ might help, but we couldn't do a log transformation of $x$ because we have 0 counts. Adding a small constant doesn't seem to help much

    ```{r, message = FALSE, warning = FALSE}
    qplot(x = log(alpha + 1), y = log(act), data = plut) +
          geom_smooth(se = FALSE)
    ```

- We could try a different transformation, like a square root:
    ```{r, message = FALSE, warning = FALSE}
    qplot(x = sqrt(alpha), y = sqrt(act), data = plut) +
          geom_smooth(se = FALSE)
    ```

    That does seem to work better.

- Let's fit using square root transformations:
    ```{r}
    plut <- mutate(plut, sq_act = sqrt(act),
                   sq_alpha = sqrt(alpha))
    lm_sq <- lm(sq_act ~ sq_alpha, data = plut)
    lm_sq
    aout <- augment(lm_sq)
    qplot(x = .fitted, y = .resid, data = aout) +
      geom_hline(yintercept = 0)
    ```
    
- That seems to work well. Let's do an overall $F$-test to see if the linear model is insufficient.

    ```{r}
    lm_full <- lm(sq_act ~ as.factor(sq_alpha), data = plut)
    anova(lm_sq, lm_full)
    ```

    This says that we do not have any evidence that the linear model is insufficient (p = 0.3).
    
- Because square roots are completely uninterpretable, I would transform back to the original scale before showing anything to my client.

- First, let's calculate prediction intervals for the range of the $x$-values. Make sure you create the *transformed* $x$-variable.

    ```{r}
    newdf <- data.frame(alpha = seq(min(plut$alpha), max(plut$alpha), length.out = 200))
    newdf <- mutate(newdf, sq_alpha = sqrt(alpha))
    predict(object = lm_sq, newdata = newdf, interval = "prediction") %>%
      cbind(newdf) ->
      newdf
    str(newdf)
    ```

- Now, let's back transform the fit and the prediction intervals.

    ```{r}
    newdf <- mutate(newdf, fit = fit^2,
                    lwr = lwr^2,
                    upr = upr^2)
    ```


- I would work hard on a visualization to my client (see STAT 412/612)

    ```{r}
    library(latex2exp)
    ggplot() +
      geom_point(data = plut, mapping = aes(x = alpha, y = act)) +
      geom_line(data = newdf, mapping = aes(x = alpha, y = fit)) +
      geom_ribbon(data = newdf, mapping = aes(x = alpha, ymin = lwr, ymax = upr), alpha = 0.2, fill = "blue") +
      ggtitle(TeX("Estimated Model: $\\sqrt{y} = 0.073 + 0.057 \\sqrt{x} + noise$")) +
      xlab("Alpha Count Rate (#/sec)") +
      ylab("pCi/g")
    ```
    
