---
title: "Linear Algebra Lab"
author: "David Gerard"
date: "`r Sys.Date()`"
output:  
  html_document:
    toc: true
    toc_depth: 4
urlcolor: "blue"
params:
  solutions: FALSE
---

```{r setup, include=FALSE}
set.seed(1)
knitr::opts_chunk$set(echo = params$solutions, 
                      eval = params$solutions, 
                      fig.align  = "center",
                      fig.height = 3)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))
```

# Blood Pressure

The following data were obtained in a study of the relation between diastolic blood pressure and age for boys 5 to 13 years old. We will manually implement a regression of blood pressure on age.

|   $i$:    | 1  | 2  | 3  | 4  | 5  | 6  | 7  |
|----------:|:--:|:--:|:--:|:--:|:--:|:--:|:--:|
| age:      | 5  | 8  | 11 | 7  | 13 | 12 | 6  |
| pressure: | 63 | 67 | 74 | 64 | 75 | 69 | 60 |

1. Create the response vector in R, call it `y`.

    ```{r}
    y <- matrix(c(63, 67, 74, 64, 75, 69, 60), ncol = 1)
    y
    ```

2. Create the design matrix in R, call it `X`.

    ```{r}
    X <- matrix(c(rep(1, 7), 5, 8, 11, 7, 13, 12, 6), ncol = 2)
    X
    ```

3. Calculate the OLS estimates from a regression of pressure on age using matrix multiplication in R. Call this vector `betahat`.

    ```{r}
    betahat <- solve(t(X) %*% X) %*% t(X) %*% y
    betahat
    ```

4. Calculate the fitted values from the above defined quantities. Call this vector `yhat`.

    ```{r}
    yhat <- X %*% betahat
    yhat
    ```

5. Calculate the residuals from the above defined quantities. Call this vector `e`.

    ```{r}
    e <- y - yhat
    e
    ```

6. Calculate the MSE from the above defined quantities. Call this value `mse`.

    ```{r}
    mse <- sum(e^2) / (length(e) - 2)
    mse
    ```

7. Calculate the standard errors of $\hat{\mathbf{\beta}}$ from the above quantities.

    ```{r}
    sqrt(mse * diag(solve(t(X) %*% X)))
    ```

8. Verify all of your calculations above using `lm()`.

    ```{r, message = FALSE}
    library(broom)
    library(dplyr)
    lmout <- lm(y ~ X[, 2])
    tidy(lmout) %>%
      select(term, estimate, std.error)
    glance(lmout)$sigma^2 ## MSE
    augment(lmout) %>%
      select(.fitted, .resid)
    ```


